Extended Data

### 2022-03-18 doc/manuscript/Dow_manuscript_Nature.Rmd
Update Dow_manuscript_Nature.Rmd

<!--on the fate of extra C-->
Our results add to a growing body of evidence for a sink limitation of stem growth[@delpierre_wood_2016;@refs], wherein global change factors known to enhance photosynthesis, such as longer growing seasons or elevated CO~2~, not cause a corresponding increase in $ANPP_{woody}$[@jiang_fate_2020;@refs].
The fate of any additional C fixed during years with warm springs remains unresolved, but possible destinations -- including respiration, NSC storage, and production of foliage, reproductive structures, roots[@zohner_how_2021], or root exudates -- generally have shorter residence times than woody growth.
Indeed, when GPP of a mature forest was increased through experimental CO~2~ enrichment, $ANPP_{woody}$ remained unchanged, while additional C was released back to the atmosphere on relatively short time scales through enhanced respiration[@jiang_fate_2020].
Consistent with this, it has been observed that C gains from an earlier spring can be offset through fall or winter respiration[@oishi_warmer_2018;@buermann_widespread_2018].
Thus, observed augmentation of NEE by warm springs[@zhou_explaining_2016;@fu_maximum_2019;@keenan_net_2014] is likely to be short lived.

@@

<!--the dominant effect of temperature on ANPP_woody is a negative effect of gs temperature-->
Regardless of the influence of spring temperatures on C cycling within the ecosystem, our results clearly demonstrate that the dominant effects of temperature on deciduous tree growth occur not in the spring, but during the peak growing season, when increased atmospheric demand associated with high temperatures limits growth[@novick_increasing_2016;@delpierre_wood_2016;@etzold_number_2022;@zweifel_why_2021]. 
Indeed, the timing of peak growth in June and July (`r EDT_phenology_parameters`, `r EDF_growth_curves_all`) coincides with the timing of the greatest sensitivity of annual growth to $T_{max}$ (Fig. `r tree_rings`, `r EDT_tree_ring_summary`).
This finding is consistent with numerous tree-ring studies demonstrating strong sensitivity of growth to drought stress or high temperatures during the peak growing season[@dorangeville_drought_2018;@helcoski_growing_2019;@anderson-teixeira_joint_2021;@babst_twentieth_2019].
<!--*It is also worth noting that growth is often influenced by temperature of the preceding year(s) through its influence on production and storage of NSCs[@refs;@helcoski], and indeed we observe many significant effects of $T_{max}$ in the preceding growing season, albeit very few in the preceding spring (`r EDF_tree_rings_TMAX_extended`).*--><!--this sentence feels tangential, but discussion of these lag effects was requested by one reviewer.-->
Warm springs may also amplify summer drought stress in some times and places, effectively canceling out any positive effects of an extended growing period[@buermann_widespread_2018;@zhang_drought_2021;@lian_summer_2022].
While such an interaction was unlikely to have played a major role within the scope of our dendrometer band study, given relatively mesic conditions and lack of significant correlation between spring temperatures and summer drought stress (see Methods), our tree-ring analysis does reveal a higher frequency of negative than positive correlations of annual growth to spring temperatures, particularly for ring porous species in cooler climates (Fig. `r tree_rings`, `r EDT_tree_ring_summary`).
Thus, warm springs can have a net negative effect on growth at some times or places[@gao_earlier_2022].

>>>>

<!--on the fate of extra C-->
Our results add to a growing body of evidence for a sink limitation of stem growth[@delpierre_wood_2016;@refs], wherein global change factors known to enhance photosynthesis, such as longer growing seasons or elevated CO~2~, do not cause a corresponding increase in $ANPP_{woody}$[@jiang_fate_2020;@refs].
The fate of any additional C fixed during years with warm springs remains unresolved, but possible destinations -- including respiration, NSC storage, and production of foliage, reproductive structures, roots[@zohner_how_2021], or root exudates -- generally have shorter residence times than woody growth.
Indeed, when GPP of a mature forest was increased through experimental CO~2~ enrichment, $ANPP_{woody}$ remained unchanged, while additional C was released back to the atmosphere on relatively short time scales through enhanced respiration[@jiang_fate_2020].
Consistent with this, it has been observed that C gains from an earlier spring can be offset through fall or winter respiration[@oishi_warmer_2018;@buermann_widespread_2018].
Thus, observed augmentation of NEE by warm springs[@zhou_explaining_2016;@fu_maximum_2019;@keenan_net_2014] is likely to be short lived.

@@

<!--the dominant effect of temperature on ANPP_woody is a negative effect of gs temperature-->
Regardless of the influence of spring temperatures on C cycling within the ecosystem, our results clearly demonstrate that the dominant effects of temperature on deciduous tree growth occur not in the spring, but during the peak growing season, when increased atmospheric demand associated with high temperatures limits growth[@novick_increasing_2016;@delpierre_wood_2016;@etzold_number_2022;@zweifel_why_2021]. 
Indeed, the timing of peak growth in June and July (`r EDT_phenology_parameters`, `r EDF_growth_curves_all`) coincides with the timing of the greatest sensitivity of annual growth to $T_{max}$ (Fig. `r tree_rings`, `r EDT_tree_ring_summary`).
This finding is consistent with numerous tree-ring studies demonstrating strong sensitivity of growth to drought stress or high temperatures during the peak growing season[@dorangeville_drought_2018;@helcoski_growing_2019;@anderson-teixeira_joint_2021;@babst_twentieth_2019].
<!--*It is also worth noting that growth is often influenced by temperature of the preceding year(s) through its influence on production and storage of NSCs[@refs;@helcoski], and indeed we observe many significant effects of $T_{max}$ in the preceding growing season, albeit very few in the preceding spring (`r EDF_tree_rings_TMAX_extended`).*--><!--this sentence feels tangential, but discussion of these lag effects was requested by one reviewer.-->
Warm springs may also amplify summer drought stress in some times and places, effectively canceling out any positive effects of an extended growing period[@buermann_widespread_2018;@zhang_drought_2021;@lian_summer_2022].
While such an interaction was unlikely to have played a major role within the scope of our dendrometer band study, given relatively mesic conditions and lack of significant correlation between spring temperatures and summer drought stress (see Methods), our tree-ring analysis does reveal a higher frequency of negative than positive correlations of annual growth to spring temperatures, particularly for ring porous species in cooler climates (Fig. `r tree_rings`, `r EDT_tree_ring_summary`).
Thus, warm springs can have a net negative effect on growth, particularly when water is limiting[@gao_earlier_2022].
